### 2019-07-11 manuscript/response.tex

\renewenvironment{quote}{\bigskip\noindent\itshape\ignorespaces}{\smallskip}

\begin{document}

\hfill \today

Dear Dr. Schoen and reviewers,

\medskip
Thank you very much for your close reading of our manuscript and supportive comments. We appreciate your constructive suggestions, each of which we address below. You will find a big difference on the number of  models  that we have fitted for the revision based on your suggestions (originally 12 now 29). The two main reasons to increase the number of models were to first address one key question by made by all referees about the symmetry of the hidden states, and second to address a remaining question about adding complexity to two-state models. For the symmetry of the hidden states  question we found that is relevant to have asymmetric hidden state rates and in our case, it strengthened the differences amongst net diversification rates. The second question,  comes from a vague idea that we had while submitting the first version of this article: by increasing from two to three states the models, are we explaining better the history of diversification? This second question could not be addressed directly by model comparison statistics because the two-state and three-state models need different input (i.e. (0,1) versus (0,1,2)). However, the species and the tree are the same and our intuition indicated that they should carry the same biological evidence despite any particular coding needed as input. During this revision, we figured a way to recode the three-state models to make them equivalent to the two-state models (the new lumped models). Thanks to this equivalency, two-state and three-state models can now be compared to argue for the importance of adding key trait information and states in diversification analyses. We also revised the original dataset thanks to the suggestion of Reviewer 1, and we were able to correct discrepancies with previous dataset as well as add more sample (651 taxa).

We hope the changes and the new models are satisfactory and exciting to you and the reviewers.

@@

We reworded this whole paragraph, as part of better explaining how polyploidization disables gametophytic SI.

\begin{quote}
--83 Nitpicky, but is it accurate to refer to this as a false-positive problem?
``False positive'' implies to me the erroneous rejection of the null, but in this case the situation is more of the null being mis-specified with respect to the question the researcher is after.
The null---no rate variation---is correctly rejected.
This section probably should include a reference to Rabosky and Goldberg.
\end{quote}

We reworded this whole paragraph, as part of better explaining the SSE family of models.

\begin{quote}
--115 I'm uncomfortable with this way of handling multiple cytotypes.
The most commonly reported cytotype could easily not be the most common cytotype in the wild (I assume that most taxa are represented by relatively few counts), and regardless, there is biology here.
Specifically, this methodology would artificially reduce the number of ploidy changes (if a taxon is polymorphic for ploidy, there's at least one event there), and the omitted changes would be disproportionately those towards the tips of the phylogeny, so not only would rates be depressed, but the bias is of a sort that the estimation of diversification parameters is particularly sensitive to (the authors will understand why this is so).
\end{quote}
We acknowledge this is a potential problem. It is true that chromosome number multiplicity can bias coding towards more (or less) transitions when it heavily depends on the most common cytotype. For the majority of species (~85\%)  the ploidy was not directly dictated by us but by the original sources of the data. In the originally sources ploidy has been classified not only most common cytotype but geographical distribution, multivalence, and/or  genome sizes. The remaining 15\% of the sample is what we assigned based on genus or family cytotype multiplicity. We cross-validated all of the assignments with the data from \citet{robertson_2011} where ploidy was assigned via a chromosome number change model using ChromEvol \citep{mayrose_2010}. We differ in exactly 35 species from \citet{robertson_2011} were we found a new reference or a new chromosome count. Classifying using a chromosome number change models only could also suffer from inflation of polyploidy events due to the lack of diploidization parameters, we found some specific species in the  \citet{robertson_2011} sample that were clearly aneuploids (one chromosome difference) that were classified as polyploids using chromosome number models, we manually corrected for those hoping to minimize obvious mistakes.

@@

\begin{quote}
--140 I lost track a little of how the data were coded -- maybe add an explicit statement as to how many tips had both forms of data, how many had only breeding system info, how many had only ploidy info (I see now, after the fact, that those data are included in the figure), and how missing data were handled
\end{quote}

Fixed.

\begin{quote}
--It's not intuitive to me what the effects of forcing the hidden state change rates to be symmetrical are.
Did the authors investigate this?
Or has it been explored in the literature?
Similarly, it seems to me that fixing the transition rates of the observed characters to be independent of the hidden state could do weird things---the point of these SSE models is that state change rates and state-dependent diversification rates need to be co-estimated
\end{quote}

We appreciate this suggestion and we followed it. The main result after fitting models with symmetric vs. asymmetric rates is that the direction  of the results does not change but the amount of difference amongst net diversification posterior distributions does. We found using Bayes factors that a model including asymmetric hidden rates is preferred over a model with equal hidden rates. Also,  the asymmetry in the trait rates have moderate to not support of being better than a model with equal trait transition rates.  %TODO put number.

@@

\begin{quote}
--221  Propagation
\end{quote}

Fixed.

\begin{quote}
--I would be more comfortable if each model was run at least twice, independently, to better assess whether the MCMC has indeed converged
\end{quote}
Unfortunately, we weren't able to run multiple chains for all models due to some limitations to our computer resources. However, we were able to test for the convergence of  three key models using Gelman's $\hat{R}$ test statistic. We tested convergence with two runs of 130,000 generations (after 5,000 generations of burn-in) for models D/P (BiSSE), D/P+A/B asym (HiSSE) and ID/CD/CP+A/B asym (MuHiSSE). The statistics as well as the Gelman plots in  \ref{figure:convergencestats} show that with that number of generations the models were converging. We increased the number of generations to be safe for all the models that we did not test and visually assessed the mixing of  models on top of demanding ESS larger than 200 for parameters.

\begin{figure}
    \centering 
      \begin{minipage}[b]{0.33\textwidth}
    \includegraphics[width=\textwidth]{Gelmanbissenodip.pdf} 
    \end{minipage}
    \begin{minipage}[b]{0.33\textwidth}
        \includegraphics[width=\textwidth]{Gelmanhissenodipasym.pdf}
        \end{minipage}
    \begin{minipage}[b]{0.33\textwidth}
             \includegraphics[width=\textwidth]{Gelmanmuhissenodipasym.pdf}
                 \end{minipage}
    \caption{ Convergence statistics and Gelman plots for D/P (BiSSE), D/P+A/B asym (HiSSE) at the top and ID/CD/CP+A/B asym (MuHiSSE,bottom). Upper limits of Gelman's test statistics should be close to 1 for an MCMC to be considered convergent.}
    \label{figure:convergencestats}
\end{figure}

\begin{quote}
--the use of run time as a description of the analysis seems odd to me.
As a reader, I'm not interested in the power of a given computer -- I want to know how many generations the MCMC was run for, what the ESSs are, etc.
\end{quote}
We ran each model for 200,000 generations. Unfortunately, we don't have access to more of 96 hours of computational time to increase those but we have calculated the convergence as indicated above.

@@

\begin{quote}
-325  These analyses (by accounting for diversification) may take care of this issue, but is there a problem with these results being driven by a non-stationary process being modeled as a stationary one?
\end{quote}

This was a point worth considering, but our model does not assume stationarity.
In Fig.~4, the contributions of each pathway are functions of time.
We can see there both the short-term and long-term behavior.
% Rosana, what do you think?- I think this is a good answer- I think it is good

@@

\begin{quote}
--beginning of the Discussion is perhaps unnecessary -- it's more like Introduction material
\end{quote}

consider % TODO

\begin{quote}
--are the results from the analyses that include diploidization coherent?
I'm not sure how to phrase this -- I guess I'm asking if there's any suggestion that that parameterization is indeed capturing the biological phenomenon of diploidization, or is it just soaking up noise?
\end{quote}

The soaking noise comes from the hidden state so in the presence of hidden states finding that diploidization matters is key. All the Bayes factors indicated strong support for models that include diploidization. However, the diploidzation parameter per se is difficult to estimate (wide credible intervals) mainly because diploidization is a genome size process that can be partially or minimally recover from ploidy data. This is the reason why we do not overstate this finding,  the importance of diploidization needs to be studied using genomes.

@@

We're pleased you found the results interesting!

\begin{quote}
I have two ``major'' comments.
The first is relatively easy to fix and is linked to the writing/structure of the paper.
I often had the impression that too much knowledge is assumed from the reader throughout the manuscript.
For instance, more information should be given in the intro regarding why self-incompatibility breakdowns with polyploidy (it is briefly explained, but only later in the manuscript).
Moreover, some methodological approaches could be better explained.
In particular, HISSE models are relatively recent and many readers may not know them well.
I thus think that it would be important to describe them better and also better explain the results (using biological terms) obtained.
I mention some examples below in the detailed notes where I think explanations could be added.
\end{quote}

We have rewritten parts of the manuscript, especially in the Introduction, to try to explain better the relationship between polyploidy and self-incompatibility, and the interpretations of hidden state models.
% E: I labeled the two paragraphs that I modified heavily.

\begin{quote}
My second comment has to do with the parameterization of the models.
Overall, I think that the authors have done a great job in justifying their models and explaining them.
I think it is important to mention.
In addition, their implementation of the models in revBayes is noteworthy and although not stated precisely in the manuscript, I believe this represents a great scientific contribution. 
Nevertheless, I question the decision to restrict the transition rates to and from the hidden states to be all equal.
Indeed, we have no idea that could make us think that this would be the case.
The authors justify this by ``Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.''
I am not completely convinced by this argument.
First, constraining transition rates can have an impact on the estimation of diversification rates.
Second, specifically because the authors are not interested in estimating the rates themselves, they should not be too worried about increasing the number of parameters.
This is especially true in a Bayesian framework where it is sometimes better to have overparameterized models than under-parameterized models (see Lemmon and Moriarty. 2004. Systematic Biology 53:265–277).
I think it would have been important to at least present the results of an unconstrained model to show that this decision to constrain the rates do not affect the conclusions.
\end{quote}
We have now fitted models with asymmetric hidden rates and asymmetric trait rates. We found that the former are important and strengthened the differences. We found no evidence that having all asymmetric rates is preferable over just asymmetric hidden for our specific example. %TODO (Figs and table)

@@

I have a few comments that I hope the authors will find useful.
\end{quote}

Thank you for this positive summary.

\begin{quote}
The first relates to the use of the hidden states in the models.
I agree completely with the rationale and meaning of including hidden states.
However, through my reading of the manuscript, I interpret the authors as using them simply to account for variation within a particular character state combination.
In other words, the authors are working under an assumption of trait-dependent diversification, with the hidden states simply providing a more nuanced view of the process.
But the authors are not considering a completely alternative view, which is that none of these trait combinations are strong drivers of diversification and that the shifts being inferred are simply coincidental.
Such an interpretation could be ruled out by explicitly fitting a model of trait-independent diversification.
This is achieved by linking the diversification rate parameters among the different character state combinations -- i.e., $ID_A=CD_A=CP_A$, and $ID_B=CD_B=CP_B$.
In other words, because the diversification rate differences are completely linked among the possible states, but unlinked by the hidden states, this forces an explicitly trait-independent model.
The fit of such a model could be included in the set of models in Table 1 and compared using Bayes Factors.
Of course, admittedly, I am not familiar with revBayes and how the model space is set up, so it's possible this issue is already being addressed.
But, if so, some clarity in the Methods/Results is warranted in how trait-independent models were evaluated (or even ruled out) in the analyses presented.
\end{quote}

We have fitted the trait independent models (character-independent CID). We appreciated this suggestion because it allowed us to parse out the effect of the traits in the diversification process. By comparing Bayes factors, models with traits are strongly preferred over CID when the breeding system is the focal trait, and there is little to no evidence that models that include ploidy are better than CID. These results strongly suggest that breeding system is much more informative than ploidy in the diversification dynamics of Solanaceae.
%TODO: write table number

@@

\begin{quote}
Along these same lines, how does revBayes deal with the symmetry problem with regards to the hidden states?
For example, imagine for one sample in the chain the speciation rate for $CP_A=0.1$ and $CP_B=0.05$, all else being equal.
Now, imagine in another sample the speciation rate parameters are exactly the same, but this time $CP_A=0.05$ and $CP_B=0.1$. 
Strangely, this would produce an identical likelihood as the previous sample (again, all else being equal), because A and B are really interchangeable.
This would, however, cause issues with how the posterior is summarized.
I should point out that nothing in the plots suggests this might be an issue, but this was something that I wondered about with the particular models used here.
\end{quote}

When we presented the first submission we were pushing the limits on implementations of diversification models in RevBayes. During the revision we found a work-around to be able to implement all the extra models we are presenting. New models with asymmetric traits are preferred and strengthen the results.

### 2019-07-11 manuscript/methods.tex
Editing more methodss

Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
% FIXME: Rosana will update numbers around here TODO
Of these, 290 had information for both states.
The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
A total of 405 taxa without any information about breeding system or ploidy were excluded.
%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
% R-This is very important but I have removed it because I think it is a full on paper by its own
%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.

@@

To test our hypotheses about lineage diversification and trait macroevolution, we fit a large suite of phylogenetic models.
All models are from the SSE framework \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
This style of model has parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).

@@

In order to investigate the association between ploidy level and diversification, we first defined a model in which taxa were classified as diploid $(D)$ or polyploid $(P)$ (M1).
This is the same model \citet{mayrose_2011} used for this question across an array of plant genera.
% R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
Although this model can be a powerful means to distinguish state-specific speciation and extinction from trait changes, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
We therefore define several models that incorporate additional forms of diversification rate heterogeneity.

Our second ploidy model includes a binary hidden trait that subdivides each observed state (M2).
In this CID model \citep{beaulieu_2016}, the hidden trait can affect diversification but the observed trait does not.
Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.

@@

Our first five breeding system models follow the same logic as the first ploidy models.
We begin with each species classified as self-incompatible $(I)$ or self-compatible $(C)$ (M11).
This is the same as the analysis of \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013}.

@@

We do not include additional models with diploidization because this reverse transition renders the models non-lumpable \citep{tarasov_2019}.  % E: further explanation needed?

@@

Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_1,\ M_0) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
There is `strong evidence'  for $M_1$ when this value is more than 2, `very strong' support when it is more than 6, and `extreme' support when it is more than 10 \citep{kass1995}.

>>>>

Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
% FIXME: Rosana will update numbers around here TODO:STILL NOT CORRECT!!!
Of these, 290 had information for both states.
The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
A total of 405 taxa without any information about breeding system or ploidy were excluded.
%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
% R-This is very important but I have removed it because I think it is a full on paper by its own
%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.

@@

In order to test our hypotheses about lineage diversification and trait macroevolution, we fitted 29 state-dependent speciation and extinction models (SSE) based on previous work from \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
SSE models parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).

@@

In order to investigate the association between ploidy level and diversification, we first defined a model in which taxa were classified as diploid $(D)$ or polyploid $(P)$ (M1).
This is the same as one of the models presented in  \citet{mayrose_2011} that was used for the exact same question across an array of plant genera.
% R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
Although this model can be a powerful means to distinguish state-specific speciation and extinction from trait changes, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
We therefore define several models that incorporate additional forms of diversification rate heterogeneity.

Our second ploidy model includes a binary hidden trait that subdivides each observed state (M2).
In this character independent model known as CID \citep{beaulieu_2016}, the hidden trait can affect diversification but the observed trait does not.
Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.

@@

Our first five breeding system models follow the same logic as the first ploidy models.
We begin with each species classified as self-incompatible $(I)$ or self-compatible $(C)$ (M11).
This is the same as the analysis of \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and larger aggregated dataset.

@@

We do not include additional models with diploidization because this reverse transition renders the models non-lumpable \citep{tarasov_2019}. The lumpability property requires that the rates going from the lumped state to the singular state are equal. Under the diploidization models CP goes back to CD with rate $\delta$ but the rate from CP back to ID is 0, and $\delta \neq 0$. Therefore, two-state models with diplodization cannot be re-coded to be compared to the tree-state models. 
% E: further explanation needed? R- I added a simple math explanation. Let me know if it is okay

@@

Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $K=ln(BF(M_1,\ M_0)) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
There is `strong evidence'  for $M_1$ when this value is more than 10, moderate support if the value is more than 1 and no evidence if the value is between -1 and 1. If the vaue of  $K$ is negative the evidence goes towards $M_0$.  \citep{kass1995}.
